Data Specialization for a Staged Lambda Calculus

### 2015-02-26 paper/splitting.tex
Checkin so I can go home. Make masking text similar to Figure 8.

\textbf{Correctness condition for \bbtwo\ term splitting.}  As
suggested informally in \ref{sec:example}, splitting a \lang\ program
(a \bbtwo\ term) should yield two monostage terms where evaluating the
second term (corresponding to second stage computations) on the output
of the first term (first stage computations), yields the same result
as direct evaluation of the \lang\ program.  More precisely, for any
$\coltwo e A$ which reduces to a residual $q$ (via $\diatwo e q$), $e$
splits (via the new relation~$\splittwosym$) into a syntactically
separated program $\pipeS p l r$ where $\reduce p b$, and $[b/l]r$ is
the same as $q$ up to an equivalence. The term $p$ generated by the
split (called the {\em precomputation}) is a monostage term
representing all the first-stage computations in $e$, and $l.r$
(called the {\em resumer}) is a monostage term representing all the
second-stage computations in $e$.  We refer to the result of the
precomputation ($b$) the {\em boundary value} of the split since it
represents the data communicated between the first and second stage
computations.  The mechanics of splitting \bbtwo\ terms and their
relationship to the outputs of evaluation is provided at right in
\ref{fig:splittingSummary}.

@@

\textbf{Correctness condition for \bbonem\ term splitting.} Defining
the correctness requirements for splitting \bbonem\ terms is
challenging because unlike \bbtwo\ terms (which reduce to monostage
residuals that can be directly compared to the resumer),
\bbonem\ terms reduce to partial values and residual tables that
represent computations at both stages.
%One complication arises from the fact that \lang\ terms at world \bbonem\ produce multistage output.
%For example, the \bbonem\ term:
%\begin{lstlisting}
%(next {1+2}, grnd{3+4})
%\end{lstlisting}
%reduces (via $\redonesym$) to the residual table and partial value:
%\begin{lstlisting}
%[yhat |-> 1+2] (next {yhat}, grnd{7})
%\end{lstlisting}

To resolve the mismatch between the multi-stage computations
represented by residual tables and partial values and the desired
monostaged outputs of splitting, we choose to define a translation,
called {\em masking}, written $\rtab \xi v \vsplito \mval i q$, which
converts residual tables and partial values into a syntactically
separated {\em masked value}.  The first component of the masked
value, $i$, is a value holding the result of all first stage
computations in $v$.  The second component, $q$, is a residual
representing all second stage computations in $\rtab \xi v$.

@@

Note that since $\coltwo e A$ reduces to an entirely second-stage
residual, its first-stage computations only exist to generate input
for second stage computations.  In contrast, since $\colmix e A$
reduces to a multi-stage residual table and partial value, its first
stage computations serve to both generate the value used as input for
the second stage {\em and also} produce a first-stage value.
Correspondingly, the precomputation $p$ generated when splitting
\bbtwo\ terms reduces only to a boundary value $b$ which is passed to
the resumer; but the combined term $c$ resulting from \bbtwo\ term
splitting reduces to a tuple containing both the first stage result
$i$ and the boundary value $b$ (motivating the name {\em combined
term}).

@@

Masking converts a partial value and associated residual table into a masked value $\mval i q$,
where $i$ is a monostage value encoding all of the first-stage information of the original partial value,
and $q$ is a monostage term encoding all of the second-stage information of the original partial value.
Remember that masking will not play a role in the final splitting algorithm, 
but instead exists to specify the contract to which splitting at \bbonem\ must adhere.
Accordingly, masking should be thought of as a definition, and as such it is in some parts arbitrary.

>>>>

\textbf{Correctness condition for \bbtwo\ term splitting.}  As
suggested informally in \ref{sec:example}, splitting a \lang\ program
(a \bbtwo\ term) should yield two monostage terms where evaluating the
second term (corresponding to second stage computations) on the output
of the first term (first stage computations), yields the same result
as direct evaluation of the original \lang\ program.  More precisely,
for any $\coltwo e A$ which reduces to a residual $q$ (via $\diatwo e
q$), $e$ splits (via the new relation~$\splittwosym$) into a
syntactically separated program $\pipeS p l r$ where $\reduce p b$,
and $[b/l]r$ is the same as $q$ up to an equivalence. The term $p$
generated by the split (called the {\em precomputation}, due to its
role producing the inputs for subsequent second stage computation) is
a monostage term representing all the first-stage computations in
$e$. The term $l.r$ (called the {\em resumer}) is a monostage term
representing all the second-stage computations in $e$.  We refer to
the result of the precomputation ($b$) the {\em boundary value} of the
split since it represents the data communicated between the first and
second stage computations.  The mechanics of splitting \bbtwo\ terms
and their relationship to the outputs of direct \lang\ evaluation is
provided at right in \ref{fig:splittingSummary}.

@@

\textbf{Correctness condition for \bbonem\ term splitting.} Defining
the correctness requirements for splitting \bbonem\ terms is
challenging because unlike \bbtwo\ terms (which reduce to monostage
residuals that can be directly compared to resumers), \bbonem\ terms
reduce to residual tables and partial values that represent
computations at both stages.
%One complication arises from the fact that \lang\ terms at world \bbonem\ produce multistage output.
%For example, the \bbonem\ term:
%\begin{lstlisting}
%(next {1+2}, grnd{3+4})
%\end{lstlisting}
%reduces (via $\redonesym$) to the residual table and partial value:
%\begin{lstlisting}
%[yhat |-> 1+2] (next {yhat}, grnd{7})
%\end{lstlisting}

To resolve the mismatch between the multi-stage computations
represented by residual tables and partial values and the desired
monostaged outputs of splitting, we choose to define a translation,
called {\em masking}, written $\rtab \xi v \vsplito \mval i q$, which
converts residual tables and partial values into a syntactically
separated {\em masked value}.  The first part of the masked value,
$i$, is a value holding the result of all first stage components of
$v$.  The second component, $q$, is a residual representing all second
stage computations in $\xi$ and $v$.

@@

Note that since $\coltwo e A$ reduces to an entirely second-stage
residual, its first-stage computations only exist to generate input
for second stage computations.  In contrast, since $\colmix e A$
reduces to a multi-stage residual table and partial value, its first
stage computations serve to both generate the value used as input for
the second stage {\em and also} produce a first-stage value.
Correspondingly, the precomputation $p$ generated when splitting
\bbtwo\ terms reduces only to a boundary value $b$ which is passed to
the resumer; but the combined term $c$ resulting from \bbonem\ term
splitting reduces to a tuple containing both the first stage result
$i$ and the boundary value $b$ (motivating the name {\em combined
term}).

@@

Masking converts a partial value and associated residual table into a
masked value $\mval i q$, where $i$ is a monostage value encoding all
first stage components of the partial value, and $q$ is a monostage
term encoding all of the second-stage computations of the partial
value.  Remember that masking will not play a role in the final
splitting algorithm, but instead exists to specify the contract to
which splitting at \bbonem\ must adhere.  Accordingly, masking should
be thought of as a definition, and as such it is in some parts
arbitrary.
